---
title: "Machine Learning and Causal Inference Final Project"
author: "Lucy Caffrey-Maffei"
date: "5/12/2022"
output: word_document
---

# Setup

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(haven)
library(splines)
library(glmnet)
library(grf)
library(recipes)
library(fastDummies)
library(haven)
library(pacman)
library(readr)
library(janitor)
# Set working directory
options(stringAsFactors=FALSE)
setwd("~/Documents/Github/Econ293_Final_Project/")
star <- read_sav("./data/STAR_Students.sav")
names(star) <- tolower(names(star))
# creating indicator variable for treatment group; 1 = in small class for at least a year in grades kinder - 3rd
star <- star %>% 
  mutate(W = if_else(gkclasstype==1 | g1classtype==1 | g2classtype==1 | g3classtype==1, 1, 0)) %>% 
  mutate(W = if_else(is.na(W), 0, W)) 
# creating indicators for what year students were in small classes. 1 = in small class, 0 = not in small class
star$gk_smallclass <- case_when(star$gkclasstype == 1 ~ 1,
                                ((star$gkclasstype == 2) | (star$gkclasstype == 3)) ~ 0)
star$g1_smallclass <- case_when(star$g1classtype == 1 ~ 1,
                                ((star$g1classtype == 2) | (star$g1classtype == 3)) ~ 0)
star$g2_smallclass <- case_when(star$g2classtype == 1 ~ 1,
                                ((star$g2classtype == 2) | (star$g2classtype == 3)) ~ 0)
star$g3_smallclass <- case_when(star$g3classtype == 1 ~ 1,
                                ((star$g3classtype == 2) | (star$g3classtype == 3)) ~ 0)
# creating school id variable for school student was in in their first year of STAR
star$schid_yr1star <- case_when(star$flagsgk == 1 ~ star$gkschid,
                                ((star$flagsgk == 0) & (star$flagsg1 == 1)) ~ star$g1schid,
                                ((star$flagsgk == 0) & (star$flagsg1 == 0) & (star$flagsg2 == 1)) ~ star$g2schid,
                                ((star$flagsgk == 0) & (star$flagsg1 == 0) & (star$flagsg2 == 0) & (star$flagsg3 == 1)) ~ star$g3schid)
# Load K-3 school level data
k3_schools <- read_sav("./data/STAR_K-3_Schools.sav")
k3_schools <- k3_schools %>% clean_names()
# create a duplicate of school data to merge into student data for each year of k-3
k3_k_schools <- k3_schools
colnames(k3_k_schools) <- paste(colnames(k3_k_schools), "k", sep = "_")
k3_g1_schools <- k3_schools
colnames(k3_g1_schools) <- paste(colnames(k3_g1_schools), "g1", sep = "_")
k3_g2_schools <- k3_schools
colnames(k3_g2_schools) <- paste(colnames(k3_g2_schools), "g2", sep = "_")
k3_g3_schools <- k3_schools
colnames(k3_g3_schools) <- paste(colnames(k3_g3_schools), "g3", sep = "_")
# Merge-in school data separately for each grade the student was in K-3 
star <- star %>% 
  left_join(k3_k_schools, by = c("gkschid" = "schid_k")) %>% 
  left_join(k3_g1_schools, by = c("g1schid" = "schid_g1")) %>% 
  left_join(k3_g2_schools, by = c("g2schid" = "schid_g2")) %>% 
  left_join(k3_g3_schools, by = c("g3schid" = "schid_g3"))
  
# Drop observations which have key variables missing
na_counts = map(star, ~sum(is.na(.)))
#SHORT TERM OUTCOMES
# Define short-term: grade 3 outcomes
short_term_outcomes <- c("g3treadss", "g3tmathss", "g3readbsraw",
                         "g3mathbsraw")
# List of variables to drop whe predicting short-term outcomes
drop_for_short_term <- c("g3tlistss", "g3sciencess", "g3socialsciss", "g3spellss",
                         "g3vocabss", "g3mathcomputss", "g3mathnumconcss", 
                         "g3mathapplss", "g3wordskillss", "g3mathbsobjraw", 
                         "g3mathbsobjpct", "g3readbsobjpct")
# Drop the other grade 3 outcomes from the list above
star_short_term <- star[,!(names(star) %in% drop_for_short_term)]
# Drop future (beyond grade 3) outcomes and characteristics
star_short_term <- star_short_term %>% 
  select(!starts_with(c("g4", "g5", "g6", "g7", "g8" , "hs")))
# LONG TERM OUTCOMES
# 7TH GRADE
# Define one long term outcome: 7th grade outcomes
g7_outcomes <- c("g7treadss", "g7tmathss")
# List of variables to drop when predicting 7th grade outcomes
drop_for_g7 <- c("g7tlangss","g7tbattss", "g7sciencess", "g7socialsciss",
                 "g7readcomprehss", "g7spellss", "g7vocabss", "g7mathcomputss",
                 "g7mathconcapplss", "g7langexpss", "g7langmechss", 
                 "g7studyskillss", "g7readbsobjraw", "g7mathbsobjraw")
# Drop the other grade 7 outcomes from the list above
star_g7 <- star[,!(names(star) %in% drop_for_g7)]
# Drop future (beyond grade 7) outcomes and characteristics)
star_g7 <- star_g7 |>
  select(!starts_with(c("g8", "hs")))
# high school GPA
hs_outcomes <- c("hsgpaoverall")
#drop other HS GPA outcomes
drop_for_hs <- c("hsgpaflang", "hsgpamath", "hsgpascience")
# Drop the other HS GPA outcomes from the list above
star_hs <- star[,!(names(star) %in% drop_for_hs)]
```

# Short-Term Effects Analysis

## 3rd grade math

### Cleaning

```{r g3 math cleaning}
```

### ATE via linear regression

```{r g3 math linear regression}
#primary analysis
bivariate_ols <- lm(g3tmathss ~ W + as.factor(gkschid)
                    + as.factor(g1schid)
                    + as.factor(g2schid)
                    + as.factor(g3schid), data=star_short_term_1)
summary(bivariate_ols)
#exploratory analyses
yearssmall_ols <- lm(g3tmathss ~ as.factor(yearssmall) 
                     + as.factor(gkschid)
                     + as.factor(g1schid)
                     + as.factor(g2schid)
                     + as.factor(g3schid), data=star_short_term_1)
summary(yearssmall_ols)
ksmall_ols <- lm(g3tmathss ~ gk_smallclass + as.factor(gkschid)
                 + as.factor(g1schid)
                 + as.factor(g2schid)
                 + as.factor(g3schid), data=star_short_term_1)
summary(ksmall_ols)
g1small_ols <- lm(g3tmathss ~ g1_smallclass 
                  + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid), data=star_short_term_1)
summary(g1small_ols)
g2small_ols <- lm(g3tmathss ~ g2_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  , data=star_short_term_1)
summary(g2small_ols)
g3small_ols <- lm(g3tmathss ~ g3_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid), data=star_short_term_1)
summary(g3small_ols)
```

### ATE via Causal Forest/AIPW

```{r g3 math aipw}
set.seed(123)
```

### HTE via Causal Forest

```{r g3 math hte}
g3_math_hte <- test_calibration(forest)
g3_math_hte
```

### Visualizations

```{r g3 math visualzation}
```

## 3rd grade reading

### Cleaning

```{r g3 reading cleaning}
```

### ATE via linear regression

```{r g3 reading linear regression}
#primary analysis
bivariate_ols <- lm(g3treadss ~ W + as.factor(gkschid)
                    + as.factor(g1schid)
                    + as.factor(g2schid)
                    + as.factor(g3schid)
                    , data=star_short_term_2)
summary(bivariate_ols)
#exploratory analyses
yearssmall_ols <- lm(g3treadss ~ as.factor(yearssmall) + as.factor(gkschid)
                     + as.factor(g1schid)
                     + as.factor(g2schid)
                     + as.factor(g3schid), data=star_short_term_2)
summary(yearssmall_ols)
ksmall_ols <- lm(g3treadss ~ gk_smallclass + as.factor(gkschid)
                 + as.factor(g1schid)
                 + as.factor(g2schid)
                 + as.factor(g3schid), data=star_short_term_2)
summary(ksmall_ols)
g1small_ols <- lm(g3treadss ~ g1_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid), data=star_short_term_2)
summary(g1small_ols)
g2small_ols <- lm(g3treadss ~ g2_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid), data=star_short_term_2)
summary(g2small_ols)
g3small_ols <- lm(g3treadss ~ g3_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid), data=star_short_term_2)
summary(g3small_ols)
```

### ATE via Causal Forest/AIPW

```{r g3 reading aipw}
set.seed(123)
```

### HTE via Causal Forest

```{r g3 reading hte}
g3_reading_hte <- test_calibration(forest)
g3_reading_hte
```

### Visualizations

```{r g3 reading visualization}
```

# Long-Term Effects Analysis

## 7th grade math

### Cleaning

```{r g7 math cleaning}
star_g7_math <- star_g7 |>
  filter(!is.na(g7tmathss))
# Define treatment and outcome vectors
W <- star_g7_math$W
outcome <- g7_outcomes[2]
Y <- star_g7_math |>
  pull(outcome) |>
  as.vector()
star_g7_x <- star_g7_math |>
  select(-W, -stdntid, -gkclasstype, -g1classtype, -g2classtype, -g3classtype, -cmpstype,-cmpsdura, -yearssmall, yearsstar, -gk_smallclass, -g1_smallclass, -g2_smallclass, -g3_smallclass) |>
  select(!starts_with(c("birth", "flag"))) |>
  select(!ends_with("schid")) |>
  select(!contains(c(g7_outcomes)))
```

### ATE via linear regression

```{r g7 math linear regression}
#primary analysis
bivariate_ols <- lm(g7tmathss ~ W 
                    + as.factor(gkschid)
                    + as.factor(g1schid)
                    + as.factor(g2schid)
                    + as.factor(g3schid)
                    + as.factor(g4schid)
                    + as.factor(g5schid)
                    + as.factor(g6schid)
                    + as.factor(g7schid), data=star_g7_math)
summary(bivariate_ols)
#other exploratory analyses
yearssmall_ols <- lm(g7tmathss ~ as.factor(yearssmall) 
                     + as.factor(gkschid)
                     + as.factor(g1schid)
                     + as.factor(g2schid)
                     + as.factor(g3schid)
                     + as.factor(g4schid)
                     + as.factor(g5schid)
                     + as.factor(g6schid)
                     + as.factor(g7schid), data=star_g7_math)
summary(yearssmall_ols)
ksmall_ols <- lm(g7tmathss ~ gk_smallclass
                 + as.factor(gkschid)
                 + as.factor(g1schid)
                 + as.factor(g2schid)
                 + as.factor(g3schid)
                 + as.factor(g4schid)
                 + as.factor(g5schid)
                 + as.factor(g6schid)
                 + as.factor(g7schid), data=star_g7_math)
summary(ksmall_ols)
g1small_ols <- lm(g7tmathss ~ g1_smallclass
                  + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  + as.factor(g4schid)
                  + as.factor(g5schid)
                  + as.factor(g6schid)
                  + as.factor(g7schid), data=star_g7_math)
summary(g1small_ols)
g2small_ols <- lm(g7tmathss ~ g2_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  + as.factor(g4schid)
                  + as.factor(g5schid)
                  + as.factor(g6schid)
                  + as.factor(g7schid)
                  , data=star_g7_math)
summary(g2small_ols)
g3small_ols <- lm(g7tmathss ~ g3_smallclass 
                  + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  + as.factor(g4schid)
                  + as.factor(g5schid)
                  + as.factor(g6schid)
                  + as.factor(g7schid), data=star_g7_math)
summary(g3small_ols)
```

### ATE via Causal Forest/AIPW

```{r g7 math aipw}
set.seed(123)
n <-nrow(star_g7_math)
# Number of rankings predictions will be ranked on
num.rankings <- 5
# Prepare for data.splitting
# Assign a fold number to each observation.
# The argument 'clusters' in the next step will mimic K-fold cross-fitting.
num.folds <- 10
folds <- sort(seq(n) %% num.folds) + 1
# Estimate a causal forest
forest <- causal_forest(X=star_g7_x, Y=Y, W=W, W.hat=mean(W),
                        clusters = star_g7_math$schid_yr1star,
                        min.node.size = 5)
#histogram of estimated HTE via causal forest
hist(predict(forest)$predictions)
#save estimates to dataframe
star_g7_math_hte <- star_g7_math
star_g7_math_hte$hte <- forest$predictions
#estimated ATE
g7_ate_math <- average_treatment_effect(forest, target.sample = "overlap", method = "AIPW")
g7_ate_math_se <- g7_ate_math[2]
g7_ate_math <- g7_ate_math[1]
print(paste0("95% CI: ", round(g7_ate_math,2),
             " +/- ", round(1.96 * g7_ate_math_se, 2)))
#important contributing variables in causal forest
var_importance <- variable_importance(forest) |> 
  as_tibble() |>
  mutate(n_col = row_number()) |>
  arrange(-V1)
important_vars_g7_math <- var_importance %>% 
  slice(1:10) %>% 
  pull(n_col)
important_vars_g7_math <- names(star_g7_x[important_vars_g7_math])
important_vars_g7_math
```

### HTE via Causal Forest

```{r g7 math hte}
g7_math_hte <- test_calibration(forest)
g7_math_hte
```

### Visualizations

```{r g7 math visualizations}
```

## 7th grade reading

### Cleaning

```{r g7 reading cleaning}
star_g7_reading <- star_g7 |>
  filter(!is.na(g7treadss))
# Define treatment and outcome vectors
W <- star_g7_reading$W
outcome <- g7_outcomes[1]
Y <- star_g7_reading |>
  pull(outcome) |>
  as.vector()
star_g7_x <- star_g7_reading |>
  select(-W, -stdntid, -gkclasstype, -g1classtype, -g2classtype, -g3classtype, -cmpstype,-cmpsdura, -yearssmall, yearsstar, -gk_smallclass, -g1_smallclass, -g2_smallclass, -g3_smallclass) |>
  select(!starts_with(c("birth", "flag"))) |>
  select(!ends_with("schid")) |>
  select(!contains(c(g7_outcomes)))
```

### ATE via linear regression

```{r g7 reading linear regression}
# primary analysis
bivariate_ols <- lm(g7treadss ~ W 
                    + as.factor(gkschid)
                    + as.factor(g1schid)
                    + as.factor(g2schid)
                    + as.factor(g3schid)
                    + as.factor(g4schid)
                    + as.factor(g5schid)
                    + as.factor(g6schid)
                    + as.factor(g7schid), data=star_g7_reading)
summary(bivariate_ols)
#exploratory analyses
yearssmall_ols <- lm(g7treadss ~ as.factor(yearssmall) + as.factor(gkschid)
                     + as.factor(g1schid)
                     + as.factor(g2schid)
                     + as.factor(g3schid)
                     + as.factor(g4schid)
                     + as.factor(g5schid)
                     + as.factor(g6schid)
                     + as.factor(g7schid), data=star_g7_reading)
summary(yearssmall_ols)
ksmall_ols <- lm(g7treadss ~ gk_smallclass + as.factor(gkschid)
                 + as.factor(g1schid)
                 + as.factor(g2schid)
                 + as.factor(g3schid)
                 + as.factor(g4schid)
                 + as.factor(g5schid)
                 + as.factor(g6schid)
                 + as.factor(g7schid), data=star_g7_reading)
summary(ksmall_ols)
g1small_ols <- lm(g7treadss ~ g1_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  + as.factor(g4schid)
                  + as.factor(g5schid)
                  + as.factor(g6schid)
                  + as.factor(g7schid), data=star_g7_reading)
summary(g1small_ols)
g2small_ols <- lm(g7treadss ~ g2_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  + as.factor(g4schid)
                  + as.factor(g5schid)
                  + as.factor(g6schid)
                  + as.factor(g7schid), data=star_g7_reading)
summary(g2small_ols)
g3small_ols <- lm(g7treadss ~ g3_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  + as.factor(g4schid)
                  + as.factor(g5schid)
                  + as.factor(g6schid)
                  + as.factor(g7schid), data=star_g7_reading)
summary(g3small_ols)
```

### ATE via Causal Forest/AIPW

```{r g7 reading aipw}
set.seed(123)
n <-nrow(star_g7_reading)
# Number of rankings predictions will be ranked on
num.rankings <- 5
# Prepare for data.splitting
# Assign a fold number to each observation.
# The argument 'clusters' in the next step will mimic K-fold cross-fitting.
num.folds <- 10
folds <- sort(seq(n) %% num.folds) + 1
# Estimate a causal forest
forest <- causal_forest(X = star_g7_x, Y=Y, W=W,W.hat=mean(W),
                        clusters = star_g7_reading$schid_yr1star,
                        min.node.size = 5)
#histogram of estimated heterogeneous effects
hist(predict(forest)$predictions)
#saving HTEs to dataframe
star_g7_reading_hte <- star_g7_reading
star_g7_reading_hte$hte <- forest$predictions
#estimated ATEs
g7_ate_reading <- average_treatment_effect(forest, target.sample = "overlap", method = "AIPW")
g7_ate_reading_se <- g7_ate_reading[2]
g7_ate_reading <- g7_ate_reading[1]
#not statistically significant
print(paste0("95% CI: ", round(g7_ate_reading,2),
             " +/- ", round(1.96 * g7_ate_reading_se, 2)))
#imortant contributing variables in causal forest
var_importance <- variable_importance(forest) |> 
  as_tibble() |>
  mutate(n_col = row_number()) |>
  arrange(-V1)
important_vars_g7_reading <- var_importance %>% 
  slice(1:10) %>% 
  pull(n_col)
important_vars_g7_reading<- names(star_g7_x[important_vars_g7_reading])
important_vars_g7_reading
```

### HTE via Causal Forest

```{r g7 reading hte}
g7_reading_hte <- test_calibration(forest)
g7_reading_hte
```

### Visualizations

```{r g7 reading visualizations}
```

## High School GPA

### Cleaning

```{r hs cleaning}
star_hs_1 <- star_hs |>
  filter(!is.na(hsgpaoverall))
# Define treatment and outcome vectors
W <- star_hs_1$W
outcome <- hs_outcomes[1]
Y <- star_hs_1 |>
  pull(outcome) |>
  as.vector()
star_hs_x <- star_hs_1 |>
  select(-W, -stdntid) |>
  select(!contains(c(hs_outcomes)))
```

### ATE via linear regression

```{r hs linear regression}
#primary analysis
bivariate_ols <- lm(hsgpaoverall ~ W + as.factor(gkschid)
                    + as.factor(g1schid)
                    + as.factor(g2schid)
                    + as.factor(g3schid)
                    + as.factor(g4schid)
                    + as.factor(g5schid)
                    + as.factor(g6schid)
                    + as.factor(g7schid)
                    + as.factor(g8schid)
                    , data=star_hs_1)
summary(bivariate_ols)
#exploratory analyses
yearssmall_ols <- lm(hsgpaoverall ~ as.factor(yearssmall) 
                     + as.factor(gkschid)
                     + as.factor(g1schid)
                     + as.factor(g2schid)
                     + as.factor(g3schid)
                     , data=star_hs_1)
summary(yearssmall_ols)
ksmall_ols <- lm(hsgpaoverall ~ gk_smallclass +as.factor(gkschid)
                 , data=star_hs_1)
summary(ksmall_ols)
g1small_ols <- lm(hsgpaoverall ~ g1_smallclass 
                  + as.factor(gkschid)
                  + as.factor(g1schid)
                  , data=star_hs_1)
summary(g1small_ols)
g2small_ols <- lm(hsgpaoverall ~ g2_smallclass+as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  , data=star_hs_1)
summary(g2small_ols)
g3small_ols <- lm(hsgpaoverall ~ g3_smallclass + as.factor(gkschid)
                  + as.factor(g1schid)
                  + as.factor(g2schid)
                  + as.factor(g3schid)
                  , data=star_hs_1)
summary(g3small_ols)
```

### ATE via Causal Forest/AIPW

```{r hs aipw}
set.seed(123)
n <- nrow(star_hs_1)
# Number of rankings predictions will be ranked on
num.rankings <- 5
# Prepare for data.splitting
# Assign a fold number to each observation.
# The argument 'clusters' in the next step will mimic K-fold cross-fitting.
num.folds <- 10
folds <- sort(seq(n) %% num.folds) + 1
# Estimate a causal forest
forest <- causal_forest(X = star_hs_x, Y=Y, W=W, W.hat=mean(W),
                        clusters = star_hs_x$hsid,
                        min.node.size = 5)
#histogram of estimated HTE via causal forest
hist(predict(forest)$predictions)
#saving HTEs to dataframe
star_hs_1_hte <- star_hs_1
star_hs_1_hte$hte <- forest$predictions
#ate predictions
hs_ate <- average_treatment_effect(forest, target.sample = "overlap", method = "AIPW")
hs_ate_se <- hs_ate[2]
hs_ate <- hs_ate[1]
#not statistically significant
print(paste0("95% CI: ", round(hs_ate,2),
             " +/- ", round(1.96 * hs_ate_se, 2)))
#important contributing variables in causal forest
var_importance <- variable_importance(forest) |> 
  as_tibble() |>
  mutate(n_col = row_number()) |>
  arrange(-V1)
important_vars_hs <- var_importance %>% 
  slice(1:10) %>% 
  pull(n_col)
important_vars_hs <- names(star_hs_x[important_vars_hs])
important_vars_hs
```

### HTE via Causal Forest

```{r hs hte}
hs_hte <- test_calibration(forest)
hs_hte
```

### Visualizations

```{r hs visualizations}
```

# Cost-Benefit Analysis

### Cleaning

```{r}
```

### Visualizations

```{r}
```

### Cleaning

```{r Collapse Data G7 Math}
library(doBy)
k3_schools_math <- summaryBy(g7tmathss ~ g3schid + W + gender + race + g3freelunch + g3tchid + g3tyears, star_g7_math)

# Define treatment and outcome vectors
W <- k3_schools_math$W
Y <- k3_schools_math$g7tmathss.mean

```

### ATE via linear regression

```{r g7 math school level linear regression}
#primary analysis
bivariate_ols <- lm(g7tmathss.mean ~ W + gender + race + g3freelunch + g3tchid + g3tyears 
                    , data=k3_schools_math)
summary(bivariate_ols)
```

### ATE via Causal Forest/AIPW

```{r g7 math aipw}
set.seed(123)
n <-nrow(k3_schools_math)
# Number of rankings predictions will be ranked on
num.rankings <- 5
# Prepare for data.splitting
# Assign a fold number to each observation.
# The argument 'clusters' in the next step will mimic K-fold cross-fitting.
num.folds <- 10
folds <- sort(seq(n) %% num.folds) + 1
# Estimate a causal forest
forest <- causal_forest(X=k3_schools_math, Y=Y, W=W, W.hat=mean(W))
#histogram of estimated HTE via causal forest
hist(predict(forest)$predictions)
#save estimates to dataframe
k3_schools_math_hte <- k3_schools_math
k3_schools_math_hte$hte <- forest$predictions
#estimated ATE
k3_schools_ate_math <- average_treatment_effect(forest, target.sample = "overlap", method = "AIPW")
k3_schools_ate_math_se <- k3_schools_ate_math[2]
k3_schools_ate_math <- k3_schools_ate_math[1]
print(paste0("95% CI: ", round(g7_ate_math,2),
             " +/- ", round(1.96 * g7_ate_math_se, 2)))
#important contributing variables in causal forest
var_importance <- variable_importance(forest) |> 
  as_tibble() |>
  mutate(n_col = row_number()) |>
  arrange(-V1)
important_vars_k3_schools_math <- var_importance %>% 
  slice(1:10) %>% 
  pull(n_col)
important_vars_k3_schools_math <- names(k3_schools_math[important_vars_k3_schools_math])
important_vars_k3_schools_math
```

### HTE via Causal Forest

```{r g7 math hte}
g7_math_hte <- test_calibration(forest)
g7_math_hte
```

### Cleaning

```{r Collapse Data G7 Reading}
library(doBy)
k3_schools_reading <- summaryBy(g7treadss ~ g3schid + W + gender + race + g3freelunch + g3tchid + g3tyears, star_g7_reading)

# Define treatment and outcome vectors
W <- k3_schools_reading$W
Y <- k3_schools_reading$g7treadss.mean

```

### ATE via linear regression

```{r g7 reading school level linear regression}
#primary analysis
bivariate_ols <- lm(g7treadss.mean ~ W + gender + race + g3freelunch + g3tchid + g3tyears 
                    , data=k3_schools_reading)
summary(bivariate_ols)
```

### ATE via Causal Forest/AIPW

```{r g7 reading school level aipw}
set.seed(123)
n <-nrow(k3_schools_reading)
# Number of rankings predictions will be ranked on
num.rankings <- 5
# Prepare for data.splitting
# Assign a fold number to each observation.
# The argument 'clusters' in the next step will mimic K-fold cross-fitting.
num.folds <- 10
folds <- sort(seq(n) %% num.folds) + 1
# Estimate a causal forest
forest <- causal_forest(X=k3_schools_reading, Y=Y, W=W, W.hat=mean(W))
#histogram of estimated HTE via causal forest
hist(predict(forest)$predictions)
#save estimates to dataframe
k3_schools_reading_hte <- k3_schools_reading
k3_schools_reading_hte$hte <- forest$predictions
#estimated ATE
k3_schools_ate_reading <- average_treatment_effect(forest, target.sample = "overlap", method = "AIPW")
k3_schools_ate_reading_se <- k3_schools_ate_reading[2]
k3_schools_ate_reading <- k3_schools_ate_reading[1]
print(paste0("95% CI: ", round(g7_ate_reading,2),
             " +/- ", round(1.96 * g7_ate_reading_se, 2)))
#important contributing variables in causal forest
var_importance <- variable_importance(forest) |> 
  as_tibble() |>
  mutate(n_col = row_number()) |>
  arrange(-V1)
important_vars_k3_schools_reading <- var_importance %>% 
  slice(1:10) %>% 
  pull(n_col)
important_vars_k3_schools_reading <- names(k3_schools_reading[important_vars_k3_schools_reading])
important_vars_k3_schools_reading
```

### HTE via Causal Forest

```{r g7 reading school level hte}
g7_reading_hte <- test_calibration(forest)
g7_reading_hte
```

### Cleaning

```{r Collapse Data G7 High School GPA}
library(doBy)
k3_schools_hsgpa <- summaryBy(hsgpaoverall ~ g3schid + W + gender + race + g3freelunch + g3tchid + g3tyears, star_hs_1)

# Define treatment and outcome vectors
W <- k3_schools_hsgpa$W
Y <- k3_schools_hsgpa$hsgpaoverall.mean

```

### ATE via linear regression

```{r g7 high school gpa school level linear regression}
#primary analysis
bivariate_ols <- lm(hsgpaoverall.mean ~ W + gender + race + g3freelunch + g3tchid + g3tyears 
                    , data=k3_schools_hsgpa)
summary(bivariate_ols)
```

### ATE via Causal Forest/AIPW

```{r g7 high school gpa school level aipw}
set.seed(123)
n <-nrow(k3_schools_hsgpa)
# Number of rankings predictions will be ranked on
num.rankings <- 5
# Prepare for data.splitting
# Assign a fold number to each observation.
# The argument 'clusters' in the next step will mimic K-fold cross-fitting.
num.folds <- 10
folds <- sort(seq(n) %% num.folds) + 1
# Estimate a causal forest
forest <- causal_forest(X=k3_schools_hsgpa, Y=Y, W=W, W.hat=mean(W))
#histogram of estimated HTE via causal forest
hist(predict(forest)$predictions)
#save estimates to dataframe
k3_schools_hsgpa_hte <- k3_schools_hsgpa
k3_schools_hsgpa_hte$hte <- forest$predictions
#estimated ATE
k3_schools_ate_hsgpa <- average_treatment_effect(forest, target.sample = "overlap", method = "AIPW")
k3_schools_ate_hsgpa_se <- k3_schools_ate_hsgpa[2]
k3_schools_hsgpa <- k3_schools_ate_hsgpa[1]
print(paste0("95% CI: ", round(g7_ate_reading,2),
             " +/- ", round(1.96 * g7_ate_reading_se, 2)))
#important contributing variables in causal forest
var_importance <- variable_importance(forest) |> 
  as_tibble() |>
  mutate(n_col = row_number()) |>
  arrange(-V1)
important_vars_k3_schools_hsgpa <- var_importance %>% 
  slice(1:10) %>% 
  pull(n_col)
important_vars_k3_schools_hsgpa <- names(k3_schools_reading[important_vars_k3_schools_hsgpa])
important_vars_k3_schools_hsgpa
```

### HTE via Causal Forest

```{r g7 reading school level hte}
hsgpa_hte <- test_calibration(forest)
hsgpa_hte
```
